---
title: "R Notebook"
output: html_notebook
---
#packages
```{r}
library(edgeR)

suppressPackageStartupMessages(library("limma"))
suppressPackageStartupMessages(library("rlist"))
#library("rstatix")
suppressPackageStartupMessages(library("broom"))
suppressPackageStartupMessages(library("ggplot2"))
library(rstatix)
library('tidyr')
library('dplyr')
```

#metadata LCL
```{r}
#which cell type to analyze?
myCell <- "LCL"
 
#read in the metadata table
metadata <-read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/LCL/metadata.txt", stringsAsFactors = FALSE)
#subtract 1 from X count to model number of additional X chromosomes
metadata$x_count <- metadata$x_count - 1

#restrict to male samples only for NPY analysis
metadata_myCell_male <- metadata[metadata$karyotype %in% c("XY") & metadata$cell_type == myCell & metadata$Technical_replicate == "N",]
rownames(metadata_myCell_male) <- metadata_myCell_male$sample

```
#X + Y chromosome genes LCL
```{r}

load('/lab/page/adrianna/RNA-seq_to_share/20221021/Autosome_log2FoldChange_per_X_or_Y_calculations/txi_lcl.rda')
txi_lcl <- as.data.frame(txi_lcl$abundance)

g_p <- read.delim("/lab/solexa_page/hannah/220516_mpra/msa/long_alignments/g_p.txt")
g_p <- g_p %>% filter(pair != "NLGN4X_NLGN4Y" & pair != "TMSB4X_TMSB4Y" & pair != "TXLNG_TXLNGY")

normCounts_expXgenes1 <- txi_lcl %>%
  dplyr::select(any_of(metadata_myCell_male$sample))


normCounts_expXgenes1 <- as.data.frame(t(as.matrix(normCounts_expXgenes1))) %>%
  dplyr::select(any_of(g_p$gene.x))


normCounts_expXgenes1$sample <- row.names(normCounts_expXgenes1)

merged_counts_by_sample <- normCounts_expXgenes1 %>% tidyr::pivot_longer(c(1:14), names_to = "gene.x", values_to = "gene_expression")
merged_counts_by_sample <- merged_counts_by_sample 
merged_counts_by_sample <- merge(merged_counts_by_sample, g_p, by = "gene.x")

write.table(x = merged_counts_by_sample, file = "/lab/solexa_page/hannah/supp_info/tables/LCL_merged_counts_by_sample.txt", quote = FALSE, sep = "\t", row.names = FALSE)

pdf(file=('/lab/solexa_page/hannah/supp_info/figures/lcl_TPM_expression_1115.pdf'), width=6, height=4)
ggplot(merged_counts_by_sample, aes(x=type_of_gene, y =  
                        gene_expression)) + 
    geom_boxplot() + 
    facet_wrap(~pair, scales = "free_y") + 
   expand_limits( y = 0) + 
    theme_pubr() + 
    ylab("TPM") + 
    geom_point() + 
    xlab("") +
  theme(strip.background =element_rect(fill="grey92")) + 
  theme(text = element_text(size = 12))  
dev.off()

merged_counts_by_sample_LCL <- merged_counts_by_sample

```

#metadata Fib
```{r}
#which cell type to analyze?
myCell <- "Fib"

#read in the metadata table
metadata <-read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/fibroblast/metadata.txt", stringsAsFactors = FALSE)
#subtract 1 from X count to model number of additional X chromosomes
metadata$x_count <- metadata$x_count - 1

#restrict to male samples only for NPY analysis
metadata_myCell_male <- metadata[metadata$karyotype %in% c("XY") & metadata$cell_type == myCell & metadata$Technical_replicate == "N",]
rownames(metadata_myCell_male) <- metadata_myCell_male$sample

```
#X + Y chromosome genes Fib
```{r}

load('/lab/page/adrianna/RNA-seq_to_share/20221021/Autosome_log2FoldChange_per_X_or_Y_calculations/txi_fib.rda')
txi_fib <- as.data.frame(txi_fib$abundance)

g_p <- read.delim("/lab/solexa_page/hannah/220516_mpra/msa/long_alignments/g_p.txt")
g_p <- g_p %>% filter(pair != "NLGN4X_NLGN4Y" & pair != "TMSB4X_TMSB4Y" & pair != "TXLNG_TXLNGY")

normCounts_expXgenes1 <- txi_fib %>%
  dplyr::select(any_of(metadata_myCell_male$sample))


normCounts_expXgenes1 <- as.data.frame(t(as.matrix(normCounts_expXgenes1))) %>%
  dplyr::select(any_of(g_p$gene.x))


normCounts_expXgenes1$sample <- row.names(normCounts_expXgenes1)

merged_counts_by_sample <- normCounts_expXgenes1 %>% tidyr::pivot_longer(c(1:14), names_to = "gene.x", values_to = "gene_expression")
merged_counts_by_sample <- merged_counts_by_sample 
merged_counts_by_sample <- merge(merged_counts_by_sample, g_p, by = "gene.x")

write.table(x = merged_counts_by_sample, file = "/lab/solexa_page/hannah/supp_info/tables/fib_merged_counts_by_sample.txt", quote = FALSE, sep = "\t", row.names = FALSE)


#pdf(file=('/lab/solexa_page/hannah/supp_info/figures/fib_TPM_expression_1115.pdf'), width=6, height=4)
ggplot(merged_counts_by_sample, aes(x=type_of_gene, y =  
                        gene_expression)) + 
    geom_boxplot() + 
    facet_wrap(~pair, scales = "free_y") + 
   expand_limits( y = 0) + 
    theme_pubr() + 
    ylab("TPM") + 
    geom_point() + 
    xlab("") +
  theme(strip.background =element_rect(fill="grey92")) + 
  theme(text = element_text(size = 12))  
#dev.off()

```

#metadata skew XXY LCL
```{r}

g_p <- read.delim("/lab/solexa_page/hannah/220516_mpra/msa/long_alignments/g_p.txt")
g_p <- g_p %>% filter(pair != "NLGN4X_NLGN4Y" & pair != "TMSB4X_TMSB4Y" & pair != "TXLNG_TXLNGY")

#which cell type to analyze?
myCell <- "LCL"

#read in the metadata table
metadata <-read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/LCL/metadata.txt", stringsAsFactors = FALSE)
#subtract 1 from X count to model number of additional X chromosomes
metadata$x_count <- metadata$x_count - 1

#skew_lines <- read.delim("skewed_lines.txt", header = FALSE) %>% filter(V1 == "LCL")
lcl_ar <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/LCL/LCL_AR.txt") %>% filter(Gene %in% c(g_p$gene.x)) %>% filter(Skewed_XCI == TRUE) #%>% filter(Sample %in% c(metadata_myCell_male$sample))
samples <- unique(lcl_ar$Sample)
values_to_remove <- c("3429_XXY", "7089B_XXY", "7194A_XXY", "7304B_XXYY", "7307A_XXY", "7367A_XX_tri21", "7369A_XX_tri21", "7370A_XX_tri21", "7372A_XX_tri21")

samples <- samples[!(samples %in% values_to_remove)]


#restrict to male samples only for NPY analysis
metadata_myCell_male <- metadata[metadata$sample %in% c(samples) & metadata$cell_type == myCell & metadata$Technical_replicate == "N",]
rownames(metadata_myCell_male) <- metadata_myCell_male$sample

```

#skew X+Y chr genes 
```{r}
load('/lab/page/adrianna/RNA-seq_to_share/20221021/Autosome_log2FoldChange_per_X_or_Y_calculations/txi_lcl.rda') #the tri 21 samples dont make it into this RDA (why?) 
txi_lcl_counts <- (txi_lcl$counts)
txi_lcl_lenght <- txi_lcl$length
rpk <- ( txi_lcl_counts*10^3 )/txi_lcl_lenght

y <- DGEList(counts=rpk)
y <- calcNormFactors(y, method = c("TMM")) #/ calculate TMM normalization factors:
# normalized counts:
cpms <- cpm(y, log=FALSE)

normCounts_expXgenes1 <- as.data.frame(cpms) %>%
  dplyr::select(any_of(metadata_myCell_male$sample))

normCounts_expXgenes1 <- as.data.frame(t(as.matrix(normCounts_expXgenes1))) %>%
  dplyr::select(any_of(g_p$gene.x))

normCounts_expXgenes1$sample <- row.names(normCounts_expXgenes1)

merged_counts_by_sample <- normCounts_expXgenes1 %>% tidyr::pivot_longer(c(1:14), names_to = "gene.x", values_to = "gene_expression")
merged_counts_by_sample <- merged_counts_by_sample 
merged_counts_by_sample <- merge(merged_counts_by_sample, g_p, by = "gene.x")

merged_counts_by_sample_LCL <- merged_counts_by_sample

```
#get Y expr LCL 
```{r} 
#which cell type to analyze?
myCell <- "LCL"

#read in the metadata table
metadata <-read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/LCL/metadata.txt", stringsAsFactors = FALSE)
#subtract 1 from X count to model number of additional X chromosomes
metadata$x_count <- metadata$x_count - 1

#restrict to male samples only for NPY analysis
metadata_myCell_male <- metadata[metadata$karyotype %in% c("XY") & metadata$cell_type == myCell & metadata$Technical_replicate == "N",]
rownames(metadata_myCell_male) <- metadata_myCell_male$sample

g_p <- read.delim("/lab/solexa_page/hannah/220516_mpra/msa/long_alignments/g_p.txt")
g_p <- g_p %>% filter(pair != "NLGN4X_NLGN4Y" & pair != "TMSB4X_TMSB4Y" & pair != "TXLNG_TXLNGY") #%>% filter(type_of_gene == "Y_gene")

normCounts_expXgenes1 <- as.data.frame(cpms) %>%
  dplyr::select(any_of(metadata_myCell_male$sample))

normCounts_expXgenes1 <- as.data.frame(t(as.matrix(normCounts_expXgenes1))) %>%
  dplyr::select(any_of(g_p$gene.x))

normCounts_expXgenes1$sample <- row.names(normCounts_expXgenes1)
merged_counts_by_sample <- normCounts_expXgenes1 %>% tidyr::pivot_longer(c(1:14), names_to = "gene.x", values_to = "gene_expression")
merged_counts_by_sample <- merged_counts_by_sample 
merged_counts_by_sample <- merge(merged_counts_by_sample, g_p, by = "gene.x")

y_gene_expr <- merged_counts_by_sample

#for tpm promoter activity of lcls 
y <- y_gene_expr %>% group_by(sample, pair) %>% mutate(ratio = gene_expression[which(type_of_gene == "Y_gene")] / gene_expression[which(type_of_gene == "X_gene")]) %>% select("pair", "ratio", "sample") %>% distinct() %>% ungroup %>% group_by(pair) %>% mutate(std = sd(ratio)) %>% mutate(num_vals = n())
write.table(x = y, file = "/lab/solexa_page/hannah/adrianna_data_linear_reg/std_LCL_ratios.txt", quote = FALSE, sep = "\t", row.names = FALSE)


y_gene_expr$name <- ifelse(y_gene_expr$type_of_gene == "X_gene", 'Xgene', 'Ygene')


y_gene_expr <- y_gene_expr %>% select("pair", "type_of_gene","name", "gene_expression")


```
#adjust by LCL AR  
```{r}
lcl_ar <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/LCL/LCL_AR.txt") %>% filter(Gene %in% c(g_p$gene.x)) %>% filter(Sample %in% c(merged_counts_by_sample_LCL$sample))

merged_counts_by_sample_LCL <- merged_counts_by_sample_LCL %>%
  rename(Gene = gene.x) %>% rename(Sample = sample)

m_lcl_ar <- merge(merged_counts_by_sample_LCL, lcl_ar, by = c("Gene", "Sample"))
m_lcl_ar$samplepair <- paste(m_lcl_ar$Sample, m_lcl_ar$pair)

merged_counts_by_sample_LCL$samplepair <- paste(merged_counts_by_sample_LCL$Sample, merged_counts_by_sample_LCL$pair)

m_lcl_ar$expr_from_Xi <- (m_lcl_ar$AR_adj * m_lcl_ar$gene_expression) / (1 + m_lcl_ar$AR_adj)

m_lcl_ar$expr_from_Xa <- m_lcl_ar$gene_expression - m_lcl_ar$expr_from_Xi
write.table(x = m_lcl_ar, file = "/lab/solexa_page/hannah/supp_info/tables/LCL_m_lcl_ar.txt", quote = FALSE, sep = "\t", row.names = FALSE)


m_lcl_ar <- m_lcl_ar %>% select(c("pair","type_of_gene", "expr_from_Xi", "expr_from_Xa")) %>% pivot_longer(cols = c(3,4), values_to = "gene_expression")

tog <- rbind(m_lcl_ar, y_gene_expr)

```
#stats and graph LCL 
```{r}


#Outliers assumption
tog %>%
  group_by(pair) %>%
  identify_outliers(gene_expression)
# Build the linear model
model  <-  lm(gene_expression ~ name*pair,
             data = tog)
# # Create a QQ plot of residuals
ggqqplot(residuals(model))
shapiro_test(residuals(model))
tog %>%
  group_by(pair) %>%
  shapiro_test(gene_expression)
ggqqplot(tog, "gene_expression", ggtheme = theme_bw()) +
facet_grid(~pair)
# #Homogeneity of variance assumption
tog %>% levene_test(gene_expression ~ name*pair)

res.kruskal <- tog  %>% group_by(pair) %>% kruskal_test(gene_expression ~ name) #all sig 

pwc <- tog  %>% group_by(pair) %>% dunn_test(gene_expression ~ name, p.adjust.method = "holm")
  
pwc <- pwc %>%  add_xy_position(x = "name",dodge = 1,step.increase = .1)

tog$name = factor(tog$name, levels=c('expr_from_Xa','Xgene', 'expr_from_Xi',  "Ygene")) 

p <- ggdotplot(tog, x = "name", y = "gene_expression",
   add = "mean_sd") + theme(legend.position = "none")

p <- facet(p, facet.by = "pair", scales = "free")

pdf(file=('/lab/solexa_page/hannah/supp_info/figures/S1_lcl_Xi_Y_2.pdf'), width=15, height=10)

p +  stat_pwc(
  tip.length = 0,
  method = "wilcox_test", label = "{p.adj.signif}",
  p.adjust.method = "bonferroni", p.adjust.by = "panel",
  hide.ns = FALSE
) 
dev.off()

write.table(x = tog, file = "/lab/solexa_page/hannah/supp_info/tables/LCL_pairs_tpm.txt", quote = FALSE, sep = "\t", row.names = FALSE)

```

#metadata skew XXY FIB
```{r}
#which cell type to analyze?
myCell <- "Fib"

#read in the metadata table
metadata <-read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/fibroblast/metadata.txt", stringsAsFactors = FALSE)
#subtract 1 from X count to model number of additional X chromosomes
metadata$x_count <- metadata$x_count - 1

#skew_lines <- read.delim("skewed_lines.txt", header = FALSE) %>% filter(V1 == "Fib")

fib_ar <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/fibroblast/Fib_AR.txt") %>% filter(Gene %in% c(g_p$gene.x)) %>% filter(Skewed_XCI == TRUE) #%>% filter(Sample %in% c(metadata_myCell_male$sample))
samples <- unique(fib_ar$Sample)


#restrict to male samples only for NPY analysis
metadata_myCell_male <- metadata[metadata$sample %in% c(samples) & metadata$cell_type == myCell & metadata$Technical_replicate == "N",]
rownames(metadata_myCell_male) <- metadata_myCell_male$sample


```

#skew X+Y chr genes 
```{r}
load('/lab/page/adrianna/RNA-seq_to_share/20221021/Autosome_log2FoldChange_per_X_or_Y_calculations/txi_fib.rda')
txi_fib_counts <- (txi_fib$counts)
txi_fib_lenght <- txi_fib$length
rpk <- ( txi_fib_counts*10^3 )/txi_fib_lenght

y <- DGEList(counts=rpk)
y <- calcNormFactors(y, method = c("TMM")) #/ calculate TMM normalization factors:
# normalized counts:
cpms <- cpm(y, log=FALSE)

g_p <- read.delim("/lab/solexa_page/hannah/220516_mpra/msa/long_alignments/g_p.txt")
g_p <- g_p %>% filter(pair != "NLGN4X_NLGN4Y" & pair != "TMSB4X_TMSB4Y" & pair != "TXLNG_TXLNGY")

normCounts_expXgenes1 <- as.data.frame(cpms) %>%
  dplyr::select(any_of(metadata_myCell_male$sample))

normCounts_expXgenes1 <- as.data.frame(t(as.matrix(normCounts_expXgenes1))) %>%
  dplyr::select(any_of(g_p$gene.x))

normCounts_expXgenes1$sample <- row.names(normCounts_expXgenes1)

merged_counts_by_sample <- normCounts_expXgenes1 %>% tidyr::pivot_longer(c(1:14), names_to = "gene.x", values_to = "gene_expression")
merged_counts_by_sample <- merged_counts_by_sample 
merged_counts_by_sample <- merge(merged_counts_by_sample, g_p, by = "gene.x")

merged_counts_by_sample_Fib <- merged_counts_by_sample

```

#get Y expr Fib
```{r}
#which cell type to analyze?
myCell <- "Fib"

#read in the metadata table
metadata <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/fibroblast/metadata.txt", stringsAsFactors = FALSE)
#subtract 1 from X count to model number of additional X chromosomes
metadata$x_count <- metadata$x_count - 1

#restrict to male samples only for NPY analysis
metadata_myCell_male <- metadata[metadata$karyotype %in% c("XY") & metadata$cell_type == myCell & metadata$Technical_replicate == "N",]
rownames(metadata_myCell_male) <- metadata_myCell_male$sample

g_p <- read.delim("/lab/solexa_page/hannah/220516_mpra/msa/long_alignments/g_p.txt")
g_p <- g_p %>% filter(pair != "NLGN4X_NLGN4Y" & pair != "TMSB4X_TMSB4Y" & pair != "TXLNG_TXLNGY") #%>% filter(type_of_gene == "Y_gene")

normCounts_expXgenes1 <- as.data.frame(cpms) %>% 
  dplyr::select(any_of(metadata_myCell_male$sample))

normCounts_expXgenes1 <- as.data.frame(t(as.matrix(normCounts_expXgenes1))) %>%
  dplyr::select(any_of(g_p$gene.x))

normCounts_expXgenes1$sample <- row.names(normCounts_expXgenes1)

merged_counts_by_sample <- normCounts_expXgenes1 %>% tidyr::pivot_longer(c(1:14), names_to = "gene.x", values_to = "gene_expression")
merged_counts_by_sample <- merged_counts_by_sample 
merged_counts_by_sample <- merge(merged_counts_by_sample, g_p, by = "gene.x")

y_gene_expr <- merged_counts_by_sample
y <- y_gene_expr %>% group_by(sample, pair) %>% mutate(ratio = gene_expression[which(type_of_gene == "Y_gene")] / gene_expression[which(type_of_gene == "X_gene")]) %>% select("pair", "ratio", "sample") %>% distinct() %>% ungroup %>% group_by(pair) %>% mutate(std = sd(ratio)) %>% mutate(num_vals = n())
write.table(x = y, file = "/lab/solexa_page/hannah/adrianna_data_linear_reg/std_fib_ratios.txt", quote = FALSE, sep = "\t", row.names = FALSE)
y_gene_expr$name <- ifelse(y_gene_expr$type_of_gene == "X_gene", 'Xgene', 'Ygene')
y_gene_expr <- y_gene_expr %>% select("pair", "type_of_gene","name", "gene_expression")

```

#adjust by Fib AR 
```{r}
lcl_ar <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/fibroblast/Fib_AR.txt") %>% filter(Gene %in% c(g_p$gene.x)) %>% filter(Sample %in% c(merged_counts_by_sample_Fib$sample)) 

merged_counts_by_sample_Fib <- merged_counts_by_sample_Fib %>%
  rename(Gene = gene.x) %>% rename(Sample = sample)

m_lcl_ar <- merge(merged_counts_by_sample_Fib, lcl_ar, by = c("Gene", "Sample"))
m_lcl_ar$samplepair <- paste(m_lcl_ar$Sample, m_lcl_ar$pair)

merged_counts_by_sample_Fib$samplepair <- paste(merged_counts_by_sample_Fib$Sample, merged_counts_by_sample_Fib$pair)

m_lcl_ar$expr_from_Xi <- (m_lcl_ar$AR_adj * m_lcl_ar$gene_expression) / (1 + m_lcl_ar$AR_adj)

m_lcl_ar$expr_from_Xa <- m_lcl_ar$gene_expression - m_lcl_ar$expr_from_Xi
write.table(x = m_lcl_ar, file = "/lab/solexa_page/hannah/supp_info/tables/Fib_m_lcl_ar.txt", quote = FALSE, sep = "\t", row.names = FALSE)

m_lcl_ar <- m_lcl_ar %>% select(c("pair","type_of_gene", "expr_from_Xi", "expr_from_Xa")) %>% pivot_longer(cols = c(3,4), values_to = "gene_expression")

tog <- rbind(m_lcl_ar, y_gene_expr)

```

#stats and graph FIB
```{r}

#Outliers assumption
tog %>%
  group_by(pair) %>%
  identify_outliers(gene_expression)
# Build the linear model
model  <-  lm(gene_expression ~ name*pair,
             data = tog)
# # Create a QQ plot of residuals
ggqqplot(residuals(model))
shapiro_test(residuals(model))
tog %>%
  group_by(pair) %>%
  shapiro_test(gene_expression)
ggqqplot(tog, "gene_expression", ggtheme = theme_bw()) +
facet_grid(~pair)
# #Homogeneity of variance assumption
tog %>% levene_test(gene_expression ~ name*pair)

res.kruskal <- tog  %>% group_by(pair) %>% kruskal_test(gene_expression ~ name) #all sig 

pwc <- tog  %>% group_by(pair) %>% dunn_test(gene_expression ~ name, p.adjust.method = "holm")
  
  
pwc <- pwc %>%  add_xy_position(x = "name",dodge = 1,step.increase = .1)

tog$name = factor(tog$name, levels=c('expr_from_Xa','Xgene', 'expr_from_Xi',  "Ygene")) 


p <- ggdotplot(tog, x = "name", y = "gene_expression", 
   add = "mean_sd") + theme(legend.position = "none")

p <- facet(p, facet.by = "pair", scales = "free")
#
pdf(file=('/lab/solexa_page/hannah/supp_info/figures/S1_fib_Xi_Y_2.pdf'), width=8, height=6)

p +  stat_pwc(
  tip.length = 0,
  method = "wilcox_test", label = "{p.adj.signif}",
  p.adjust.method = "bonferroni", p.adjust.by = "panel",
  hide.ns = FALSE
)
dev.off()

write.table(x = tog, file = "/lab/solexa_page/hannah/supp_info/tables/Fib_pairs_TPM.txt", quote = FALSE, sep = "\t", row.names = FALSE)

```
#TPM  promoter activity LCL
```{r}
g_p <- read.delim("/lab/solexa_page/hannah/220516_mpra/msa/long_alignments/g_p.txt")
g_p <- g_p %>% filter(pair != "NLGN4X_NLGN4Y" & pair != "TMSB4X_TMSB4Y" & pair != "TXLNG_TXLNGY")

lcl <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/LCL_pairs_tpm.txt") #copy over to tables

lcl <- lcl %>%
  pivot_wider(
    id_cols = c(pair, type_of_gene),
    names_from = name,
    values_from = gene_expression, 
    values_fn = {mean}
  )

acusm_LCL <- read.delim("/lab/solexa_page/hannah/supp_info/tables/aucsm_lcl.txt") 

join_max_a <- full_join(lcl, acusm_LCL, by = c("pair", "type_of_gene"))

xg <- join_max_a[1:7,c(1:5,7:9)]
xg$areas_x <- xg$areas

xy <- join_max_a[8:14,c(1,6:9)]
xy$areas_y <- xy$areas

together <- merge(xg,xy, by = 'pair')

together$Y_X <- together$Ygene / together$Xgene

together$areaY_areaX <- together$areas_y / together$areas_x
together$Y_Xi <- together$Ygene / together$expr_from_Xi
together$Y_Xa <- together$Ygene / together$expr_from_Xa

std_lcl <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/std_LCL_ratios.txt") %>% select("std", "pair") %>% distinct()

together <- left_join(together, std_lcl, by = "pair")



pdf(file=('/lab/solexa_page/hannah/supp_info/figures/S16_LCL_area_expr_diff.pdf'), width=2.3, height=2.3)
ggplot(together, aes(x=areaY_areaX , y = Y_X)) +   geom_point() + 
 # xlim(0,100) +
  theme_pubr() + 
  stat_cor(method = "spearman", label.x = 0, label.y = 0.8) + 
  #geom_smooth(method='lm', formula= y~x, se= FALSE) + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  geom_abline(intercept = 1, slope = 0) + 
  geom_vline(xintercept = 1) + 
  geom_label_repel(aes(label = pair)) + 
  ylab('Y TPM / X TPM') + 
  xlab('Y total activity / X total activity') + 
  ylim(0,1.6) +
  xlim(0,1.6) + 
  scale_x_continuous(limits = c(0, 1.6), expand = expansion(mult = c(0, 0))) + 
  scale_y_continuous(limits = c(0, 1.6), expand = expansion(mult = c(0, 0))) + 

  geom_errorbar(aes(ymin = Y_X - std, ymax = Y_X + std), width = 0.01 )
dev.off()

```
#TPM  promoter activity FIB
```{r}
g_p <- read.delim("/lab/solexa_page/hannah/220516_mpra/msa/long_alignments/g_p.txt")
g_p <- g_p %>% filter(pair != "NLGN4X_NLGN4Y" & pair != "TMSB4X_TMSB4Y" & pair != "TXLNG_TXLNGY")

lcl <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/Fib_pairs_tpm.txt")
#lcl$gene_expression <- log2(lcl$gene_expression)
lcl <- lcl %>%
  pivot_wider(
    id_cols = c(pair, type_of_gene),
    names_from = name,
    values_from = gene_expression, 
    values_fn = {median}
  )


acusm_LCL <- read.delim("/lab/solexa_page/hannah/supp_info/tables/aucsm_full_fib.txt") 

join_max_a <- full_join(lcl, acusm_LCL, by = c("pair", "type_of_gene"))

xg <- join_max_a[1:7,c(1:5,7:9)]
xg$areas_x <- xg$areas

xy <- join_max_a[8:14,c(1,6:9)]
xy$areas_y <- xy$areas

together <- merge(xg,xy, by = 'pair')
together$Y_X <- together$Ygene / together$Xgene
together$areaY_areaX <- together$areas_y / together$areas_x
together$Y_Xi <- together$Ygene / together$expr_from_Xi
together$Y_Xa <- together$Ygene / together$expr_from_Xa

std_fib <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/std_fib_ratios.txt") %>% select("std", "pair") %>% distinct()

together <- left_join(together, std_fib, by = "pair")

pdf(file=('/lab/solexa_page/hannah/supp_info/figures/5E_fib_area_expr_diff1.pdf'), width=2.3, height=2.3)

ggplot(together, aes(x=areaY_areaX , y = Y_X)) +   geom_point() + 
 # xlim(0,100) +
  theme_pubr() + 
  stat_cor(method = "spearman", label.x = 0, label.y = 0.8) + 
  #geom_smooth(method='lm', formula= y~x, se= FALSE) + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  geom_label_repel(aes(label = pair)) + 
    geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  geom_abline(intercept = 1, slope = 0) + 
  geom_vline(xintercept = 1) + 
  ylab('Y TPM / X TPM') + 
  xlab('Y total activity / X total activity') + 
  ylim(0,1.3) +
  xlim(0,1.3) + 
    scale_x_continuous(limits = c(0, 1.3), expand = expansion(mult = c(0, 0))) + 
  scale_y_continuous(limits = c(0, 1.3), expand = expansion(mult = c(0, 0))) + 

  geom_errorbar(aes(ymin = Y_X - std, ymax = Y_X + std), width = 0.01)

dev.off()

```

#ratio comparisons  fib
```{r}
auc_fib <- read.delim("/lab/solexa_page/hannah/supp_info/tables/aucsm_full_fib.txt")
xy_counts <- read.delim('/lab/solexa_page/hannah/adrianna_data_linear_reg/std_fib_ratios.txt') %>% select(c("pair", "ratio"))
xy_counts$type <- "YtoX"                                                                                                          #geTMM
xaxi_counts <- read.delim("/lab/solexa_page/hannah/supp_info/tables/Fib_m_lcl_ar.txt") #geTMM


auc_fib %<>% group_by(pair) %>% mutate(ratio = areas[which(type_of_gene == "Y_gene")] / (areas[which(type_of_gene == "X_gene")] )) %>% filter(type_of_gene == "X_gene") %>% select(c("pair", "ratio"))

xaxi_counts$ratio <- xaxi_counts$expr_from_Xi / (xaxi_counts$expr_from_Xa)

xaxi_counts <- xaxi_counts %>% select(c("pair",'ratio'))
xaxi_counts$type <- "XitoXa"                                                                                                          #geTMM

xaxiy <- rbind(xaxi_counts,xy_counts)

xaxiy %>% ggplot(aes(x=pair,y=ratio, color = type)) + 
  geom_boxplot() + 
  theme_pubr() + 
      theme(axis.text.y = element_text(family = "Helvetica"),axis.text.x = element_text(family = "Helvetica", angle = 90, vjust = 0.5, hjust = 1)) 


# xaxi_counts %<>% select(c("pair", "xaxi_ratio")) %>% group_by(pair) %>% mutate(avg_xaxi_ratio = median(xaxi_ratio)) %>% select(c("pair", "avg_xaxi_ratio")) %>% distinct()

# merge_r <- merge(xaxi_counts, auc_fib, by = "pair")

# ggplot(merge_r, aes(x=avg_xaxi_ratio, y = ratio )) + 
#   geom_point() + 
#   theme_pubr()+ geom_smooth(method='lm', formula= y~x, se= FALSE) + stat_regline_equation(label.y = 1)

# mer_xyc_r <- merge(xy_counts, merge_r, by = 'pair')
# 
# mer_xyc_r %>% filter(sample == "7169C_XY" | sample == "7276D_XY") %>%
# ggplot(aes(x=xy_ratio, y = ratio )) + 
#   geom_point() + 
#   theme_pubr() + 
#   ylim(0.4,1) + 
#   xlim(0.4,1) + geom_smooth(method='lm', formula= y~x, se= FALSE) + stat_regline_equation(label.y = 1) 
# 
# new_df <-  mer_xyc_r %>% select(c("pair", "xy_ratio")) %>% group_by(pair) %>% mutate(avg_xay_ratio = median(xy_ratio)) %>% select(c("pair", "avg_xay_ratio")) %>% distinct()

# mergeddf <- merge(merge_r, new_df, by = "pair") 

``` 
#ratio comparisons LCL
```{r}
auc_fib <-  read.delim("/lab/solexa_page/hannah/supp_info/tables/aucsm_lcl.txt")
xy_counts <- read.delim('/lab/solexa_page/hannah/adrianna_data_linear_reg/std_LCL_ratios.txt') %>% select(c("pair", "ratio"))
xy_counts$type <- "YtoX"                                                                              #geTMM
xaxi_counts <- read.delim("/lab/solexa_page/hannah/supp_info/tables/LCL_m_lcl_ar.txt")                #geTMM


auc_fib %<>% group_by(pair) %>% mutate(ratio = areas[which(type_of_gene == "Y_gene")] / (areas[which(type_of_gene == "X_gene")] )) %>% filter(type_of_gene == "X_gene") %>% select(c("pair", "ratio"))

xaxi_counts$ratio <- xaxi_counts$expr_from_Xi / (xaxi_counts$expr_from_Xa)

xaxi_counts <- xaxi_counts %>% select(c("pair",'ratio'))
xaxi_counts$type <- "XitoXa"                                                                                                          #geTMM

xaxiy <- rbind(xaxi_counts,xy_counts)

xaxiy %>% ggplot(aes(x=pair,y=ratio, color = type)) + 
  geom_boxplot() + 
  theme_pubr() + 
      theme(axis.text.y = element_text(family = "Helvetica"),axis.text.x = element_text(family = "Helvetica", angle = 90, vjust = 0.5, hjust = 1)) 


# auc_fib <- read.delim("/lab/solexa_page/hannah/supp_info/tables/aucsm_lcl.txt")
# xy_counts <- read.delim('/lab/solexa_page/hannah/supp_info/tables/std_LCL_ratios.txt')
# xaxi_counts <- read.delim("/lab/solexa_page/hannah/supp_info/tables/LCL_m_lcl_ar.txt"
# )
# 
# auc_fib %<>% group_by(pair) %>% mutate(ratio = areas[which(type_of_gene == "X_gene")] / (areas[which(type_of_gene == "X_gene")] + areas[which(type_of_gene == "Y_gene")])) %>% filter(type_of_gene == "X_gene") %>% select(c("pair", "ratio"))
# 
# 
# xy_counts %<>% group_by(sample, pair) %>% mutate(xy_ratio = gene_expression[which(type_of_gene == "X_gene")] / (gene_expression[which(type_of_gene == "X_gene")] + gene_expression[which(type_of_gene == "Y_gene")])) %>% filter(type_of_gene == "X_gene")
# 
# xaxi_counts$xaxi_ratio <- xaxi_counts$expr_from_Xa / (xaxi_counts$expr_from_Xa + xaxi_counts$expr_from_Xi)
# 
# xaxi_counts %<>% select(c("pair", "xaxi_ratio")) %>% group_by(pair) %>% mutate(avg_xaxi_ratio = median(xaxi_ratio)) %>% select(c("pair", "avg_xaxi_ratio")) %>% distinct()
# 
# merge_r <- merge(xaxi_counts, auc_fib, by = "pair")
# 
# ggplot(merge_r, aes(x=avg_xaxi_ratio, y = ratio )) + 
#   geom_point() + 
#   theme_pubr()+ geom_smooth(method='lm', formula= y~x, se= FALSE) + stat_regline_equation(label.y = 1)
# 
# mer_xyc_r <- merge(xy_counts, merge_r, by = 'pair')
# 
# mer_xyc_r %>% filter(sample == "7365A_XY") %>% 
# ggplot(aes(x=xy_ratio, y = ratio )) + 
#   geom_point() + 
#   theme_pubr() + 
#   ylim(0.4,1) + 
#   xlim(0.4,1) + geom_smooth(method='lm', formula= y~x, se= FALSE) + stat_regline_equation(label.y = 1) 
# 
# new_df <-  mer_xyc_r %>% select(c("pair", "xy_ratio")) %>% group_by(pair) %>% mutate(avg_xay_ratio = median(xy_ratio)) %>% select(c("pair", "avg_xay_ratio")) %>% distinct()
# 
# mergeddf <- merge(merge_r, new_df, by = "pair") 

``` 

#plot ratios of YvX expression in TPMS 
```{r}

fib_ratios <- read.delim('/lab/solexa_page/hannah/adrianna_data_linear_reg/std_fib_ratios.txt') #GeTMM 

# pdf(file=('/lab/solexa_page/hannah/supp_info/figures/poster_fib_ratio_figure.pdf'), width=5, height=3)

ggplot(fib_ratios, aes(x=pair,y = ratio)) + 
  geom_boxplot() + 
  geom_point() + 
  theme_pubr()
# dev.off()


lcl_ratios <- read.delim("/lab/solexa_page/hannah/adrianna_data_linear_reg/std_LCL_ratios.txt") 

# pdf(file=('/lab/solexa_page/hannah/supp_info/figures/poster_lcl_ratio_figure.pdf'), width=5, height=3)
ggplot(lcl_ratios, aes(x=pair,y = ratio)) + 
  geom_boxplot() + 
  geom_point() + 
  theme_pubr()
# dev.off()


```








